---
title: "Project 1 Group 5"
author: "Aaron Zalki, Mia Siracusa, Lidiia Tronina, John Suh, Henry Vasquez"
date: "6/21/2020"
---

```{r}

library(dplyr)
data <- readxl::read_excel('./Data Set for Class.xls')%>%
 mutate(SeriesInd = as.Date(SeriesInd, "1899-12-30"))

data%>%
  summary()

```


```{r}
data1 <- data %>%
  filter(between(SeriesInd, as.Date("2011-05-06"),as.Date("2017-10-13")))
data1%>%
  summary()
```
```{r}
data1%>%
  filter(!complete.cases(.)) 

```

```{r}
data_cc <- data1%>%
  filter(complete.cases(.)) 
medians <- data_cc %>%
  group_by(group) %>%
  summarise(med = median(Var01))
data1[is.na(data1$Var01) & data1$group == 'S01', ]$Var01 <- medians$med[medians$group == 'S01'][1] 
data1[is.na(data1$Var01) & data1$group == 'S02', ]$Var01 <- medians$med[medians$group == 'S02'][1] 
data1[is.na(data1$Var01) & data1$group == 'S03', ]$Var01 <- medians$med[medians$group == 'S03'][1] 
data1[is.na(data1$Var01) & data1$group == 'S04', ]$Var01 <- medians$med[medians$group == 'S04'][1] 
data1[is.na(data1$Var01) & data1$group == 'S05', ]$Var01 <- medians$med[medians$group == 'S05'][1] 
data1[is.na(data1$Var01) & data1$group == 'S06', ]$Var01 <- medians$med[medians$group == 'S06'][1] 

medians <- data_cc %>%
  group_by(group) %>%
  summarise(med = median(Var02))
data1[is.na(data1$Var02) & data1$group == 'S05', ]$Var02 <- medians$med[medians$group == 'S05'][1] 
data1[is.na(data1$Var02) & data1$group == 'S06', ]$Var02 <- medians$med[medians$group == 'S06'][1] 

medians <- data_cc %>%
  group_by(group) %>%
  summarise(med = median(Var03))
data1[is.na(data1$Var03) & data1$group == 'S01', ]$Var03 <- medians$med[medians$group == 'S01'][1] 
data1[is.na(data1$Var03) & data1$group == 'S02', ]$Var03 <- medians$med[medians$group == 'S02'][1] 
data1[is.na(data1$Var03) & data1$group == 'S03', ]$Var03 <- medians$med[medians$group == 'S03'][1] 
data1[is.na(data1$Var03) & data1$group == 'S04', ]$Var03 <- medians$med[medians$group == 'S04'][1] 
data1[is.na(data1$Var03) & data1$group == 'S05', ]$Var03 <- medians$med[medians$group == 'S05'][1] 
data1[is.na(data1$Var03) & data1$group == 'S06', ]$Var03 <- medians$med[medians$group == 'S06'][1] 


medians <- data_cc %>%
  group_by(group) %>%
  summarise(med = median(Var05))
data1[is.na(data1$Var05) & data1$group == 'S01', ]$Var05 <- medians$med[medians$group == 'S01'][1] 
data1[is.na(data1$Var05) & data1$group == 'S02', ]$Var05 <- medians$med[medians$group == 'S02'][1] 
data1[is.na(data1$Var05) & data1$group == 'S03', ]$Var05 <- medians$med[medians$group == 'S03'][1] 
data1[is.na(data1$Var05) & data1$group == 'S04', ]$Var05 <- medians$med[medians$group == 'S04'][1] 
data1[is.na(data1$Var05) & data1$group == 'S05', ]$Var05 <- medians$med[medians$group == 'S05'][1] 
data1[is.na(data1$Var05) & data1$group == 'S06', ]$Var05 <- medians$med[medians$group == 'S06'][1] 

medians <- data_cc %>%
  group_by(group) %>%
  summarise(med = median(Var07))
data1[is.na(data1$Var07) & data1$group == 'S01', ]$Var07 <- medians$med[medians$group == 'S01'][1] 
data1[is.na(data1$Var07) & data1$group == 'S02', ]$Var07 <- medians$med[medians$group == 'S02'][1] 
data1[is.na(data1$Var07) & data1$group == 'S03', ]$Var07 <- medians$med[medians$group == 'S03'][1] 
data1[is.na(data1$Var07) & data1$group == 'S04', ]$Var07 <- medians$med[medians$group == 'S04'][1] 
data1[is.na(data1$Var07) & data1$group == 'S05', ]$Var07 <- medians$med[medians$group == 'S05'][1] 
data1[is.na(data1$Var07) & data1$group == 'S06', ]$Var07 <- medians$med[medians$group == 'S06'][1] 

data1%>%
  summary()

```
S01 – Forecast  Var01, Var02
S02 – Forecast  Var02, Var03
S03 – Forecast  Var05, Var07
S04 – Forecast  Var01, Var02
S05 – Forecast  Var02, Var03
S06 – Forecast  Var05, Var07



##S01 – Forecast  Var01

##S01 – Forecast  Var02

##S02 – Forecast  Var02

##S02 – Forecast  Var03

##S03 – Forecast  Var05

##S03 – Forecast  Var07

##SUBSET DATA BY GROUP VARIABLES

```{r}
s01_var01 <- data1%>%
  filter(group =='S01' ) %>%
  select(SeriesInd, Var01)

s01_var02 <- data1%>%
  filter(group =='S01' ) %>%
  select(SeriesInd, Var02)

s02_var02 <- data1%>%
  filter(group =='S02' ) %>%
  select(SeriesInd, Var02)

s02_var03 <- data1%>%
  filter(group =='S02' ) %>%
  select(SeriesInd, Var03)

s03_var05 <- data1%>%
  filter(group =='S03' ) %>%
  select(SeriesInd, Var05)

s03_var07 <- data1%>%
  filter(group =='S03' ) %>%
  select(SeriesInd, Var07)
```

#PLOT TIME SERIES

```{r}
library(fpp2)
s01_var01<- ts(s01_var01[,2], start = c(2011,88), frequency = 261)
autoplot(s01_var01)

s01_var02<- ts(s01_var02[,2], start = c(2011,88), frequency = 261)
autoplot(s01_var02)

s02_var02<- ts(s02_var02[,2], start = c(2011,88), frequency = 261)
autoplot(s02_var02)

s02_var03<- ts(s02_var03[,2], start = c(2011,88), frequency = 261)
autoplot(s02_var03)

s03_var05<- ts(s03_var05[,2], start = c(2011,88), frequency = 261)
autoplot(s03_var05)

s03_var07<- ts(s03_var07[,2], start = c(2011,88), frequency = 261)
autoplot(s03_var07)

```


#FIND THE OUTLIERS

```{r}
s01_var01outlier <- which.max(s01_var01)
s01_var01outlier
s01_var02outlier <- which.max(s01_var02)
s01_var02outlier
s02_var02outlier <- which.max(s02_var02)
s02_var02outlier
s02_var03outlier <- which.max(s02_var03)
s02_var03outlier
s03_var05outlier <- which.max(s03_var05)
s03_var05outlier
s03_var07outlier <- which.max(s03_var07)
s03_var07outlier
```

##LOOK AT SEASONAL FREQUENCIES

```{r}
frequency(s01_var01)
frequency(s01_var02)
frequency(s02_var02)
frequency(s02_var03)
frequency(s03_var05)
frequency(s03_var07)
```

##SEASONAL PLOTS

```{r}
ggseasonplot(s01_var01)
ggseasonplot(s01_var02)
ggseasonplot(s02_var02)
ggseasonplot(s02_var03)
ggseasonplot(s03_var05)
ggseasonplot(s03_var07)
```

##SUBSERIES PLOT

```{r}
ggsubseriesplot(s01_var01)
ggsubseriesplot(s01_var02)
ggsubseriesplot(s02_var02)
ggsubseriesplot(s02_var03)
ggsubseriesplot(s03_var05)
ggsubseriesplot(s03_var07)
```


##CREATE ACF PLOTS

```{r}
ggAcf(s01_var01)
ggAcf(s01_var02)
ggAcf(s02_var02)
ggAcf(s02_var03)
ggAcf(s03_var05)
ggAcf(s03_var07)
```

##Autocorrelation example-ASK Lidia what is this?

```{r}
autoplot(s01_var01)
ggAcf(s01_var01)
maxlag_s01_var01<-5
```


##WHITE NOISE TESTING

```{r}
autoplot(s01_var01)
autoplot(diff(s01_var01))
ggAcf(diff(s01_var01))
Box.test(diff(s01_var01), lag=522, type="Ljung")

autoplot(s01_var02)
autoplot(diff(s01_var02))
ggAcf(diff(s01_var02))
Box.test(diff(s01_var02), lag=522, type="Ljung")

autoplot(s02_var02)
autoplot(diff(s02_var02))
ggAcf(diff(s02_var02))
Box.test(diff(s02_var02), lag=522, type="Ljung")

autoplot(s02_var03)
autoplot(diff(s02_var03))
ggAcf(diff(s02_var03))
Box.test(diff(s02_var03), lag=522, type="Ljung")

autoplot(s03_var05)
autoplot(diff(s03_var05))
ggAcf(diff(s03_var05))
Box.test(diff(s03_var05), lag=522, type="Ljung")

autoplot(s03_var07)
autoplot(diff(s03_var07))
ggAcf(diff(s03_var07))
Box.test(diff(s03_var07), lag=522, type="Ljung")
```

##FORECASTS USING NAIVE AND SNAIVE

```{r}
s01_var01n<-naive(s01_var01,140)
s01_var02n<-naive(s01_var02,140)
s02_var02n<-naive(s02_var02,140)
s02_var03n<-naive(s02_var03,140)
s03_var05n<-naive(s03_var05,140)
s03_var07n<-naive(s03_var07,140)
autoplot(s01_var01n)
summary(s01_var01n)
autoplot(s01_var02n)
summary(s01_var02n)
autoplot(s02_var02n)
summary(s02_var02n)
autoplot(s02_var03n)
summary(s02_var03n)
autoplot(s03_var05n)
summary(s03_var05n)
autoplot(s03_var07n)
summary(s03_var07n)
s01_var01sn<-snaive(s01_var01,140)
s01_var02sn<-snaive(s01_var02,140)
s02_var02sn<-snaive(s02_var02,140)
s02_var03sn<-snaive(s02_var03,140)
s03_var05sn<-snaive(s03_var05,140)
s03_var07sn<-snaive(s03_var07,140)
autoplot(s01_var01sn)
summary(s01_var01sn)
autoplot(s01_var02sn)
summary(s01_var02sn)
autoplot(s02_var02sn)
summary(s02_var02sn)
autoplot(s02_var03sn)
summary(s02_var03sn)
autoplot(s03_var05sn)
summary(s03_var05sn)
autoplot(s03_var07sn)
summary(s03_var07sn)

```

##FITTED VALUES AND RESIDUALS

```{r}
checkresiduals(s01_var01n)
checkresiduals(s01_var02n)
checkresiduals(s02_var02n)
checkresiduals(s02_var03n)
checkresiduals(s03_var05n)
checkresiduals(s03_var07n)
checkresiduals(s01_var01sn)
checkresiduals(s01_var02sn)
checkresiduals(s02_var02sn)
checkresiduals(s02_var03sn)
checkresiduals(s03_var05sn)
checkresiduals(s03_var07sn)
```


##TRAINING AND TEST SETS NON SEASONAL

```{r}
train <- subset(s01_var01, end = 1296)
naive_fc <- naive(train, h = 324)
mean_fc <- meanf(train, h = 108)
accuracy(naive_fc, s01_var01)
accuracy(mean_fc, s01_var01)
autoplot(naive_fc)
autoplot(mean_fc)
#naive better

train <- subset(s01_var02, end = 1296)
naive_fc <- naive(train, h = 324)
mean_fc <- meanf(train, h = 108)
accuracy(naive_fc, s01_var02)
accuracy(mean_fc, s01_var02)
autoplot(naive_fc)
autoplot(mean_fc)
#naive better

train <- subset(s02_var02, end = 1296)
naive_fc <- naive(train, h = 324)
mean_fc <- meanf(train, h = 108)
accuracy(naive_fc, s02_var02)
accuracy(mean_fc, s02_var02)
autoplot(naive_fc)
autoplot(mean_fc)
#naive better

train <- subset(s02_var03, end = 1296)
naive_fc <- naive(train, h = 324)
mean_fc <- meanf(train, h = 108)
accuracy(naive_fc, s02_var03)
accuracy(mean_fc, s02_var03)
autoplot(naive_fc)
autoplot(mean_fc)
#mean better

train <- subset(s03_var05, end = 1296)
naive_fc <- naive(train, h = 324)
mean_fc <- meanf(train, h = 108)
accuracy(naive_fc, s03_var05)
accuracy(mean_fc, s03_var05)
autoplot(naive_fc)
autoplot(mean_fc)
#naive better

train <- subset(s03_var07, end = 1296)
naive_fc <- naive(train, h = 324)
mean_fc <- meanf(train, h = 108)
accuracy(naive_fc, s03_var07)
accuracy(mean_fc, s03_var07)
autoplot(naive_fc)
autoplot(mean_fc)
#naive better


```

##TRAINING AND TEST SETS SEASONAL

```{r}
train1 <- window(s01_var01, end = c(2011, 133))
train2 <- window(s01_var01, end = c(2012, 203))
train3 <- window(s01_var01, end = c(2013, 201))

# Produce forecasts using snaive()
fc1 <- snaive(train1, h = 34)
fc2 <- snaive(train2, h = 51)
fc3 <- snaive(train3, h = 51)

# Use accuracy() to compare the MAPE of each series
accuracy(fc1, s01_var01)["Test set", "MAPE"]
accuracy(fc2, s01_var01)["Test set", "MAPE"]
accuracy(fc3, s01_var01)["Test set", "MAPE"]

autoplot(fc1)
autoplot(fc2)
autoplot(fc3)

# fc1 had lowest
```


##Exponential Smoothing


```{r}
fcs01_var01 <- ses(s01_var01, h = 140)
summary(fcs01_var01)
autoplot(fcs01_var01)
autoplot(fcs01_var01) + autolayer(fitted(fcs01_var01))

fcs01_var02 <- ses(s01_var02, h = 140)
summary(fcs01_var02)
autoplot(fcs01_var02)
autoplot(fcs01_var02) + autolayer(fitted(fcs01_var02))

fcs02_var02 <- ses(s02_var02, h = 140)
summary(fcs02_var02)
autoplot(fcs02_var02)
autoplot(fcs02_var02) + autolayer(fitted(fcs01_var01))

fcs02_var03 <- ses(s02_var03, h = 140)
summary(fcs02_var03)
autoplot(fcs02_var03)
autoplot(fcs02_var03) + autolayer(fitted(fcs02_var03))

fcs03_var05 <- ses(s03_var05, h = 140)
summary(fcs03_var05)
autoplot(fcs03_var05)
autoplot(fcs03_var05) + autolayer(fitted(fcs03_var05))

fcs03_var07 <- ses(s03_var07, h = 140)
summary(fcs03_var07)
autoplot(fcs03_var07)
autoplot(fcs03_var07) + autolayer(fitted(fcs03_var07))


```

##SES VS NAIVE

```{r}
train <- subset(s01_var01, end = length(s01_var01) - 324)
fcses <- ses(train, h = 324)
fcnaive <- naive(train, h = 324)
accuracy(fcses, s01_var01)
accuracy(fcnaive, s01_var01)
#No difference

train <- subset(s01_var02, end = length(s01_var02) - 324)
fcses <- ses(train, h = 324)
fcnaive <- naive(train, h = 324)
accuracy(fcses, s01_var02)
accuracy(fcnaive, s01_var02)
#SES better

train <- subset(s02_var02, end = length(s02_var02) - 324)
fcses <- ses(train, h = 324)
fcnaive <- naive(train, h = 324)
accuracy(fcses, s02_var02)
accuracy(fcnaive, s02_var02)
#naive better

train <- subset(s02_var03, end = length(s02_var03) - 324)
fcses <- ses(train, h = 324)
fcnaive <- naive(train, h = 324)
accuracy(fcses, s02_var03)
accuracy(fcnaive, s02_var03)
#No difference

train <- subset(s03_var05, end = length(s03_var05) - 324)
fcses <- ses(train, h = 324)
fcnaive <- naive(train, h = 324)
accuracy(fcses, s03_var05)
accuracy(fcnaive, s03_var05)
#naive better

train <- subset(s03_var07, end = length(s03_var07) - 324)
fcses <- ses(train, h = 324)
fcnaive <- naive(train, h = 324)
accuracy(fcses, s03_var07)
accuracy(fcnaive, s03_var07)
#No difference

```

##Exponential Smoothing with Trend

```{r}
fcholt <- holt(s01_var01, h = 140)
summary(fcholt)
autoplot(fcholt)
checkresiduals(fcholt)

fcholt <- holt(s01_var02, h = 140)
summary(fcholt)
autoplot(fcholt)
checkresiduals(fcholt)

fcholt <- holt(s02_var02, h = 140)
summary(fcholt)
autoplot(fcholt)
checkresiduals(fcholt)

fcholt <- holt(s02_var03, h = 140)
summary(fcholt)
autoplot(fcholt)
checkresiduals(fcholt)

fcholt <- holt(s03_var05, h = 140)
summary(fcholt)
autoplot(fcholt)
checkresiduals(fcholt)

fcholt <- holt(s03_var07, h = 140)
summary(fcholt)
autoplot(fcholt)
checkresiduals(fcholt)
```


##Exponential smoothing with trend and seasonality-getting error

```{r}
.f = function() {train <- subset(s01_var01, end = length(s01_var01) - 324)
fchw <- hw(train, seasonal = "additive", h = 324)
fcsn <- snaive(train, h = 324)
accuracy(fchw, s01_var01)
accuracy(fcsn, s01_var01)
autoplot(fchw)

train <- subset(s01_var02, end = length(s01_var02) - 324)
fchw <- hw(train, seasonal = "additive", h = 324)
fcsn <- snaive(train, h = 324)
accuracy(fchw, s01_var02)
accuracy(fcsn, s01_var02)
autoplot(fchw)

train <- subset(s02_var02, end = length(s02_var02) - 324)
fchw <- hw(train, seasonal = "additive", h = 324)
fcsn <- snaive(train, h = 324)
accuracy(fchw, s02_var02)
accuracy(fcsn, s02_var02)
autoplot(fchw)

train <- subset(s02_var03, end = length(s02_var03) - 324)
fchw <- hw(train, seasonal = "additive", h = 324)
fcsn <- snaive(train, h = 324)
accuracy(fchw, s02_var03)
accuracy(fcsn, s02_var03)
autoplot(fchw)

train <- subset(s03_var05, end = length(s03_var05) - 324)
fchw <- hw(train, seasonal = "additive", h = 324)
fcsn <- snaive(train, h = 324)
accuracy(fchw, s03_var05)
accuracy(fcsn, s03_var05)
autoplot(fchw)

train <- subset(s03_var07, end = length(s03_var07) - 324)
fchw <- hw(train, seasonal = "additive", h = 324)
fcsn <- snaive(train, h = 324)
accuracy(fchw, s03_var07)
accuracy(fcsn, s03_var07)
autoplot(fchw)}

```


##AUTOMATIC MODEL SELECTION Exponential Smoothing ets (error,trend,seasonality)

```{r}
#Remember, a model passes the Ljung-Box test when the p-value is greater than 0.05. 

fit <- ets(s01_var01)
checkresiduals(fit)
autoplot(forecast(fit))

fit2 <- stlf(s01_var01)
checkresiduals(fit2)
autoplot(forecast(fit2))


fit <- ets(s01_var02)
checkresiduals(fit)
autoplot(forecast(fit))

fit2 <- stlf(s01_var02)
checkresiduals(fit2)
autoplot(forecast(fit2))


fit <- ets(s02_var02)
checkresiduals(fit)
autoplot(forecast(fit))

fit2 <- stlf(s02_var02)
checkresiduals(fit2)
autoplot(forecast(fit2))

fit <- ets(s02_var03)
checkresiduals(fit)
autoplot(forecast(fit))

fit2 <- stlf(s02_var03)
checkresiduals(fit2)
autoplot(forecast(fit2))

fit <- ets(s03_var05)
checkresiduals(fit)
autoplot(forecast(fit))

fit2 <- stlf(s03_var05)
checkresiduals(fit2)
autoplot(forecast(fit2))

fit <- ets(s03_var07)
checkresiduals(fit)
autoplot(forecast(fit))

fit2 <- stlf(s03_var07)
checkresiduals(fit2)
autoplot(forecast(fit2))



```



```{r}
s01_var01 %>% decompose(type = "multiplicative") %>%
  autoplot()  + xlab("Year")
```